Reproducible Research: Peer Assessment 1
=============================
##Loading and processing the data  

So, we are going to load activity data via read.csv 

```{r}
activity <- read.csv("activity.csv")
```
## No let's calculate mean total number of steps taken per day?

We calculate the number of steps taken per day by using tapply.

```{r}
step_sum <- tapply(activity$steps, activity$date, sum, na.rm = TRUE, simplify = TRUE)
```
The next step is to make a histogram of the total number of steps taken each day.

```{r}
hist(step_sum, xlab = "number of steps", main = "Histogram of the total number of steps taken each day")
```

The second step is to calculate the mean and median total number of steps taken per day.

```{r}
mean(step_sum)
median(step_sum)
```

## What is the average daily activity pattern?

```{r}
activity$interval <- as.factor(activity$interval)
step_average <- tapply(activity$steps, activity$interval, sum, na.rm = TRUE, 
    simplify = TRUE)/length(levels(activity$date))
```

Then, we make a time series plot of the average number of steps taken.

```{r}
plot(x = levels(activity$interval), y = step_average, type = "l", xlab = "time", 
    ylab = "number of steps", main = "Average number of steps taken in 5-minute interval across all days")
```

We can find the 5-minute interval which contains the maximum number of steps on average across all the days.

```{r}
names(which.max(step_average))
```

The maximum number of steps on average across all the days is 104 and its corresponding 5-minute interval is 835

##Imputing missing values

At the beginning, we calculate the total number of missing values in the dataset.

```{r}
na_number <- sum(!complete.cases(activity$steps))
```

The total number of missing values in the dataset is 2304.

Then we use the mean/median of that 5-minute interval to fill in all of the missing values in the dataset. The original dataset with the missing data is stored in **original_data** and the clean dataset with the missing data filled in is stored in **clean_data**.
```{r}
original_data <- activity
for (i in 1:length(activity$steps)) {
    if (is.na(activity$steps[i])) {
        for (j in 1:length(step_average)) {
            if (as.character(activity$interval[i]) == names(step_average[j])) 
                activity$steps[i] = step_average[j]
        }
    }
}
clean_data <- activity
```

After that, we make a histogram of the total number of steps taken each day for the clean dataset.
```{r}
clean_step_sum <- tapply(clean_data$steps, clean_data$date, sum, na.rm = TRUE, 
    simplify = TRUE)
hist(clean_step_sum, xlab = "number of steps", main = "Histogram of the total number of steps taken each day")
```

We calculate the mean and median total number of steps taken per day for the clean dataset.
```{r}
mean(clean_step_sum)
median(clean_step_sum)
```

The mean and median total number of steps taken per day are 1.0581 ?? 104 and 1.0395 ?? 104. Compared with the estimates from the first part of the assignment, the difference of mean and median value are -1226.7842 and 0. We find that the shape of two histograms change a little, especially for the number of steps below 10000. The median value have no difference from the estimates from the first part of the assignment. However, the mean value increase significantly. After imputation it's expected that estimators increase, the probable reason for no change in median is due to missing values were imputed by averages of the variable. 

## Are there differences in activity patterns between weekdays and weekends?

At first, we create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
Sys.setlocale("LC_TIME", "en")
```
```{r}
weekday <- weekdays(as.Date(clean_data$date, "%Y-%m-%d"))
for (i in 1:length(weekday)) {
    if ((weekday[i] == "Saturday") | (weekday[i] == "Sunday")) 
        weekday[i] = "weekend" else weekday[i] = "weekday"
}
clean_data$weekday <- as.factor(weekday)
```

Then we make a panel plot containing a time series plot of the average number of of steps taken in each 5-minute interval across all weekday days or weekend days.

```{r}
clean_data2 <- split(clean_data, clean_data$weekday)
weekday <- clean_data2$weekday
weekend <- clean_data2$weekend
step_average_weekday <- tapply(weekday$steps, weekday$interval, sum, simplify = TRUE)/(length(weekday$weekday)/288)
step_average_weekend <- tapply(weekend$steps, weekend$interval, sum, simplify = TRUE)/(length(weekend$weekday)/288)
output <- data.frame(steps = c(step_average_weekday, step_average_weekend), 
    interval = c(levels(activity$interval), levels(activity$interval)), weekday = as.factor(c(rep("weekday", 
        length(step_average_weekday)), rep("weekend", length(step_average_weekend)))))
library(lattice)
xyplot(steps ~ interval | weekday, data = output, layout = c(1, 2), ylab = "number of steps", 
    main = "Average number of steps for all weekday days or weekend days")
```




